SeedLing: Building and using a seed corpus \\ for the Human Language Project

### 2014-04-28 docs/computEL-final/clean.tex
Added reference to Xia Lewis Lewis paper, tidied some discussion

\paragraph{UDHR.} Although there is a pre-compiled version of the UDHR data in the Natural Language ToolKit (NLTK) corpora distribution,\footnote{\url{http://nltk.googlecode.com/svn/trunk/nltk_data/index.xml}}
the distribution is laden with encoding problems. Instead, we used the plaintext files available from the Unicode website,\footnote{\url{http://unicode.org/udhr/d}} which are free from encoding issues. The first four lines of each file record metadata, and the rest is the translation of the UDHR. This dataset is extremely clean, and simply required segmentation into sentences.

\paragraph{Wikipedia.}
One major issue with using the Wikipedia dump is the problem of separating text from abundant source-specific markup. To convert compressed Wikipedia dumps to textfiles, we used the WikiExtractor\footnote{\url{http://medialab.di.unipi.it/wiki/Wikipedia_Extractor}} tool. After conversion into textfiles, we used several regular expressions to delete residual Wikipedia markup and so-called ``magic words"\footnote{\url{http://en.wikipedia.org/wiki/Help:Magic_words}}.

@@

\subsection{Language Codes}
We use ISO~639-3 as our standard set of language codes, since it aims for universal coverage, and has widespread acceptance in the community. % As a result, we needed to make sure that all our data was labelled with these codes.  
Data from ODIN and the UDHR already used this standard, and hence did not pose problems.

@@

Omniglot does not give codes at all, but only the language name. To resolve this issue, we automatically converted language names to codes using information from the SIL website.\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}}

@@

It is important to make sure that the data we have compiled will be available to future researchers, regardless of how the surrounding infrastructure changes. \newcite{bird2003port} describe a set of best practices for maintaining portability of digital information, outlining seven dimensions along which this can vary. Following this advice, we have ensured that all our data is available as plain text files, with utf-8 encoding, labelled with the relevant ISO 639-3 code. Metadata is stored separately. We have written an API to allow access to the data according to the guidelines of \newcite{abney2010universal}, who remain agnostic as to the specific form of data storage. If, for reasons of space or speed, an alternative format would be preferred, the data would be straightfoward to convert since it can be accessed according to these guidelines.

>>>>

\paragraph{UDHR.} We used the plaintext UDHR files available from the Unicode website\footnote{\url{http://unicode.org/udhr/d}} which uses UTF-8 encoding for all languages. The first four lines of each file record metadata, and the rest is the translation of the UDHR. This dataset is extremely clean, and simply required segmentation into sentences.

\paragraph{Wikipedia.}
One major issue with using the Wikipedia dump is the problem of separating text from abundant source-specific markup. To convert compressed Wikipedia dumps to textfiles, we used the WikiExtractor\footnote{\url{http://medialab.di.unipi.it/wiki/Wikipedia_Extractor}} tool. After conversion into textfiles, we used several regular expressions to delete residual Wikipedia markup and so-called ``\emph{magic words}".\footnote{\url{http://en.wikipedia.org/wiki/Help:Magic_words}}

@@

\subsection{Language Codes}
As \newcite{xia2010multilingual} explain, language names do not always suffice to identify languages, since many names are ambiguous. For this reason, sets of language codes exist to more accurately identify languages. We use ISO~639-3\footnote{url{http://www-01.sil.org/iso639-3/default.asp}} as our standard set of codes, since it aims for universal coverage, and has widespread acceptance in the community. The data from ODIN and the UDHR already used this standard.

To facilitate the standardization of language codes, we have written a python API that can be used to query information about a language or a code, fetching up-to-date information from SIL International (which maintains the ISO~639-3 code set), as well as from Ethnologue.

@@

Omniglot does not give codes at all, but only the language name. To resolve this issue, we automatically converted language names to codes using information from the SIL website.

@@

It is important to make sure that the data we have compiled will be available to future researchers, regardless of how the surrounding infrastructure changes. \newcite{bird2003port} describe a set of best practices for maintaining portability of digital information, outlining seven dimensions along which this can vary. Following this advice, we have ensured that all our data is available as plain-text files, with UTF-8 encoding, labelled with the relevant ISO 639-3 code. Metadata is stored separately. This allows users to easily process the data using the programming language or software of their choice.

To allow access to the data following Abney and Bird's guidelines, as discussed in section \ref{sec:data}, we have written an API, which we distribute along with the data.  Abney and Bird remain agnostic to the specific file format used, but if an alternative format would be preferred, the data would be straightfoward to convert since it can be accessed according to these guidelines.  As examples of functionality, our API allows a user to fetch all sentences in a given language, all sentences from a given source.
